---
title: "Worksheet 4c"
author: "Carl"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg
```{r}


```
1a. Show your solutions on how to import a csv file into the environmen
```{r}
mpg <- read.csv("mpg.csv")


```

1b. Which variables from mpg dataset are categorical?
```{r}
# all the colums exept colum "X"

```

1c. Which are continuous variables?
```{r}

mpg$model
mpg$displ
mpg$cyl
mpg$cty
mpg$hwy

```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
maxmod <- table(mpg$manufacturer)
#dodge has most model

maxvarofmod <- table(mpg$model)
#caravan 2wd has the most Variations


```
2a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}

tableofmm <- data.frame(mpg$manufacturer, mpg$model)

unofmm <- unique(tableofmm)

ddofun <- data.frame(unofmm[,0],unofmm[,1])

```

2b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}


```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```