---
title: "Worksheet 4c"
author: "Carl"
date: "2023-11-22"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg
```{r}


```
1a. Show your solutions on how to import a csv file into the environmen
```{r}
mpg <- read.csv("mpg.csv")


```

1b. Which variables from mpg dataset are categorical?
```{r}

varCat <- mpg[,c(2:4,6:12)]
varCat

```

1c. Which are continuous variables?
```{r}

carCon <- mpg[,c(4:6,9,10)]
carCon

```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
maxmod <- table(mpg$manufacturer)
maxmod
#dodge has most model


maxvarofmod <- table(mpg$model)
maxvarofmod
#caravan 2wd has the most Variations


```


2a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
library(dplyr)

manufacturer_groups <- mpg %>%
  group_by(mpg$manufacturer) %>%
  summarise(unique_models = toString(unique(model)))
manufacturer_groups

manufacturer_counts <- mpg %>%
  group_by(manufacturer) %>%
  summarise(unique_model_count = n_distinct(model))
manufacturer_counts

```

2b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}

par(mfrow = c(1,2))

barplot(manufacturer_counts$unique_model_count, names.arg = manufacturer_counts$manufacturer,
        col = "violet", border = "black", main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", ylab = "Number of Unique Models",
        las = 2)

library(ggplot2)

ggplotofUMM <- ggplot(manufacturer_counts, aes(x = manufacturer, y = unique_model_count)) +
  geom_bar(stat = "identity", fill = "violet", color = "black") +
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal()
ggplotofUMM

par(mfrow = c(1,1))
 
```

3. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
```{r}

manufacturer_model_counts <- mpg %>%
  group_by(mpg$manufacturer, mpg$model) %>%
  summarise(count = n(), .groups = "drop")

manufacturer_model_counts

ggplot(manufacturer_counts, aes(x = reorder(manufacturer, -unique_model_count), y = unique_model_count)) +
  geom_bar(stat = "identity", fill = "violet", color = "black") +
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

3a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}

ggplot(mpg, aes(model, manufacturer)) +
  geom_count()

```

3b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}

# not really. I would add color

ggplot(mpg, aes(model, manufacturer, color = factor(cyl))) +
  geom_count()

```

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}

top_20 <- head(mpg, 20)

ggplot(top_20, aes(x = interaction(model, year), fill = interaction(model, year))) +
  geom_bar() +
  labs(title = "Count of Model-Year Combinations (Top 20 Observations)",
       x = "Model-Year Combination",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}

carsPerModel <- mpg %>%
  group_by(model) %>%
  summarise(number_of_cars = n())

carsPerModel

```

4a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.
```{r}

top20 <- head(carsPerModel, 20)

plotofTop20 <- ggplot(top20, aes(x = model, y = number_of_cars, fill = model)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Number of Cars per Model", 
       x = "Model" , 
       y = "Number of cars") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

plotofTop20


```

4b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}


lotofTop20 <- ggplot(top20, aes(x = model, y = number_of_cars, fill = model)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Number of Cars per Model", 
       x = "Model" , 
       y = "Number of cars") +
  coord_flip()

print(lotofTop20)


```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.
```{r}

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  scale_color_gradient(low = "violet", high = "darkviolet") +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal()


```
5a. How would you describe its relationship? Show the codes and its result
```{r}
# number of cylenders increases the displacement also increases

```

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?
```{r}

ggplot(mpg, aes(x = displ, y = hwy, color = year)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG (Mapped with Year)",
       x = "Engine Displacement",
       y = "Highway MPG") +
  theme_minimal() +
  scale_color_gradient(low = "blue", high = "red")


```


6. Import the traffic.csv onto your R environment.
a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.
```{r}
library(readr)
traffic <- read_csv("traffic.csv")


varoftraffic <- colnames(traffic)

lengthVT <- length(varoftraffic)

# Number of Variables is 4

```
6b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}

selected_junctions <- c(1, 2, 3, 4)

junction_data <- traffic[traffic$Junction %in% selected_junctions, ]
junction_data
```


6c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)

ggplot(traffic, aes(x = DateTime, y = Vehicles, color = as.factor(Junction))) +
  geom_line() +
  scale_color_manual(values = c("red", "green", "blue", "violet")) +  # Adjust with your colors
  labs(title = "Traffic by Junction",
       x = "Date Time",
       y = "Number of Vehicles") +
  theme_minimal() +
  facet_wrap(~Junction, scales = "free_y", ncol = 1) +
  guides(color = guide_legend(title = "Junction"))

```

7. From alexa_file.xlsx, import it to your environment
```{r}
library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")

```

7a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.
```{r}

obsofalex <- nrow(alexa_file)
obsofalex

colofalex <- ncol(alexa_file)
colofalex

cat("Number of Observations:", obsofalex, "\n")
cat("Number of Columns:", colofalex, "\n")

```
7b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
library(dplyr)

groupofVar <- alexa_file %>%
  group_by(variation) %>%
  summarise(total_count = n())
groupofVar

```
7c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.
```{r}

result <- alexa_file %>%
  group_by(variation) %>%
  summarise(total_count = n())

plotofVar <- ggplot(result, aes(x = variation, y = total_count, fill = variation)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = rainbow(n = nrow(result))) +  # Rainbow colors
  labs(title = "Counts of Variations",
       x = "Variations",
       y = "Total Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plotofVar

```
7d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.
```{r}

alexa_file$date <- as.Date(alexa_file$date)

alexa_file$month <- format(alexa_file$date, "%m")

countsss <- alexa_file %>%
  group_by(month) %>%
  summarise(num_reviews = n())

monthly_review_counts <- table(countsss)

ggplot(countsss, aes(x = month, y = num_reviews, group = 1)) +
  geom_line(color = "blue") +
  labs(title = "Number of Verified Reviews Per Month",
       x = "Month of 2017", y = "Number of Reviews")

```

7e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}

ggplot(alexa_file, aes(x = variation, y = rating, fill = variation)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_manual(values = rainbow(n = length(unique(alexa_file$variation)))) +  # Rainbow colors
  labs(title = "Relationship Between Variations and Ratings",
       x = "Variations",
       y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

variation_ratings <- alexa_file %>%
  group_by(variation) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE))

topvar <- variation_ratings %>%
  filter(avg_rating == max(avg_rating)) %>%
  pull(variation)

cat("Variation with the Highest Average Rating:", topvar, "\n")


```
